### 2024-02-12 parts/appendix.tex
big big rewrite

\begin{proof}[Proof of \cref{decidable-commutative-rat:lemma}]
    Because $f$ is a \kl{$\Rel$-rational series},
    there exists a $n \in \Nat$,
    vectors $v_1$ and $v_2$,
    and $n$ by $n$ integer matrices $\seqof{M_a}{a \in \Sigma}$,
    such that
    \begin{equation*}
        \forall w \in \Sigma^*,
        f(w) = {}^t v_1 \left(\prod_{i = 1}^{|w|} M_{w_i}\right) v_2 \quad .
    \end{equation*}
    To simplify notations, 
    let us write $M_w \defined \prod_{i = 1}^{|w|} M_{w_i}$,
    and 
    $S \defined \setof{M_w}{w \in \Sigma^+}$ be the
    semigroup generated by $\seqof{M_a}{a \in \Sigma}$.

@@

\begin{proof}[Proof of
    \cref{pre-compose-growth-commut:lemma}]
    If $f \in \ZPoly[k]$, then 
    for every \kl{polyregular function} $g \in \Poly[\ell]$,
    $f \circ g \in \ZPoly[k+ \ell - 1]$
    \cite[]{LOPEZ23b}.

@@

@@

@@

@@

@@

@@

\begin{proof}[Proof of
    \cref{q-o-prefix-cool:fact}]
    Let us write $Q_i$ and $O_i$ for the value of the variables
    $Q$ and $O$ at step $i$ of the \texttt{while loop}.
    We prove the desired property by induction on $i$.

@@

    If the \texttt{else} branch was taken, then there exists $u \in O_i$ such
    that $Q_{i+1} = Q_i \cup \set{u}$ and $O_{i+1} = O_i \setminus \set{ u }
    \cup \setof{ ua }{ a \in \Sigma}$. We conclude that $Q_{i+1} \cup O_{i+1} =
    Q_i \cup O_i \cup \setof{ ua }{a \in \Sigma}$ continues to be \kl{downwards
    closed} for $\prefleq$. Let $v \in Q_{i+1} \cup O_{i+1}$ be such that $ua
    \prefleq v$ for some $a \in \Sigma$. Then $u \prefleq v$, and $u = v$ since
    $u$ was a maximal element. As a consequence, $ua$ is a maximal element for
    all $a \in \Sigma$. Assume by contradiction that $ua$ is comparable with
    some $v \in O_{i+1}$ with $ua \neq v$, it cannot be that $ua \prefleq v$ by
    the above argument, and if $v \prefleq ua$ with $v \neq ua$, then $v
    \prefleq u$ and $u = v$, which is absurd since $v \not \in O{i+1}$.
    We have concluded that $O_{i+1}$ continues to have pairwise incomparable
    elements.
\end{proof}
\begin{proof}[Proof of \cref{wqo-implies-termination:lemma}]
    Assume towards a contradiction that
    \cref{residual:algo} does not terminate.
    Then, the \texttt{else} branch in the \texttt{while loop}
    must be taken infinitely often.
    This means that the set $Q$ of states grows arbitrarily large.

    Let us write $\seqof{Q_i}{i \in \Nat}$ for the set of states $Q$ at step
    $i$ of the execution of \cref{residual:algo}. Applying
    \cref{q-o-prefix-cool:fact}, we know that for all $i \in \Nat$, $Q_i$ is
    \kl{downwards closed} for $\prefleq$. Let us write $Q_\infty \defined
    \bigcup_{i \in \Nat} Q_i$. The set $Q_\infty$ is infinite, and is
    \kl{downwards closed} for $\prefleq$. As a consequence, it is an infinite
    tree with a finite branching (at most $\card{\Sigma}$), and has an infinite
    branch $\seqof{u_j}{j \in \Nat}$ thanks to König’s lemma.

@@

    As a consequence, if $\MSO^q(u) \leq \MSO^q(v)$, then 
    for all $w \in \Sigma^*$: 
    \begin{align*}
        & f(vw) - f(uw) \\
        &= 
        \sum_{i = 1}^n m_i
        \left[
            \vcount{\phi_i(\vec{x})} (vw) -
            \vcount{\phi_i(\vec{x})} (uw)
        \right] \\
        &= 
        \sum_{i = 1}^n
        m_i
            \sum_{\phi_i \in t(\vec{x})}
        \left[
            \vcount{t(\vec{x})}(vw)
            -
            \vcount{t(\vec{x})}(uw)
        \right] \\
        &= 
        \sum_{i = 1}^n
        m_i
        \sum_{1 \leq j \leq N_0}
        \sum_{\phi_i \in t_l^j(\vec{y}) \odot t_r^j(\vec{z})}
        \\
        &\quad\quad
        \underbrace{
        \left[
            \vcount{t_r^j(\vec{y})}(v)
            -
            \vcount{t_r^j(\vec{y})}(u)
        \right] 
    }_{ \in \Nat }
            \times 
            \vcount{t_l^j(\vec{z})}(w)
    \end{align*}

@@

>>>>

\begin{proof}[Proof of \cref{size-of-word-nsf:ex}]
    Let us define $M \defined (\set{1}, \times)$ which is 
    a finite \kl{aperiodic monoid}, $\mu \colon \Sigma^* \to M$
    defined by $\mu(w) \defined 1$, and
    $\pi \colon M^2 \to \Nat$
    that is the constant function equal to $1$.
    We check that for all $w \in \Sigma^*$:
    \begin{equation*}
        \pi^\dagger(w)
        =
        \sum_{uv = w} 1
        =
        \card{w} + 1
        = f(w)
        \quad . 
        \qedhere
    \end{equation*}
\end{proof}

\begin{proof}[Proof of \cref{decidable-commutative-poly:lemma}, in the polyregular case]
    Let $\Sigma$ be a finite alphabet, endowed with a
    total ordering over its letters,
    and let $f \colon \Sigma^* \to \Rel$ be a
    \kl{$\Rel$-polyregular function}.
    The map $\polysort \colon \Sigma^* \to \Sigma^*$
    that sorts the letters in the input word $w$ according
    to the chosen ordering is a \kl{polyregular function}
    \cite[Proposition II.12]{LOPEZ23b}.
    As a consequence,
    $(f \circ \polysort) \in \ZPoly$.
    Finally, $f$ is \kl{commutative} if and only
    if $f = f \circ \polysort$, which is decidable
    in the case of $\ZPoly$
    \cite[Corollary II.24]{LOPEZ23b}, and more generally in the case 
    of $\ZRat$ \cite[Corollary 3.6 p 38]{berstel2011noncommutative}.
\end{proof}

\begin{proof}[Proof of \cref{decidable-commutative-rat:lemma} in the general case]
    Because $f$ is a \kl{$\Rel$-rational series},
    there exists a $n \in \Nat$,
    vectors $v_1$ and $v_2$,
    and $n$ by $n$ integer matrices $\seqof{M_a}{a \in \Sigma}$,
    such that
    \begin{equation*}
        \forall w \in \Sigma^*,
        f(w) = {}^t v_1 \left(\prod_{i = 1}^{|w|} M_{w_i}\right) v_2 \quad .
    \end{equation*}
    To simplify notations, 
    let us write $M_w \defined \prod_{i = 1}^{|w|} M_{w_i}$,
    and 
    $S \defined \setof{M_w}{w \in \Sigma^+}$ be the
    semigroup generated by $\seqof{M_a}{a \in \Sigma}$.

@@

\begin{proof}[Proof of
    \cref{pre-compose-growth-commut:lemma}]
    If $f \in \ZPoly[k]$, then 
    for every \kl{polyregular function} $g \in \Poly[\ell]$,
    $f \circ g \in \ZPoly[k+ \ell - 1]$
    \cite{LOPEZ23b}.

@@

\begin{proof}[Proof of \cref{negative-not-nrat:ex}]
    The function $w \mapsto |w|$ is a \kl{$\Nat$-polyregular function}.
    Thus, 
    $P(X) \defined X$ is
    a \kl{$\Nat$-rational polynomial}. Similarly,
    $w \mapsto |w|^2 + 3$ is a \kl{$\Nat$-polyregular function},
    showing that $Q(X) \defined X^2 + 3$
    is a \kl{$\Nat$-rational polynomial}.
    Finally, 
    $T(X) \defined - X$ cannot be 
    a \kl{$\Nat$-rational polynomial} as \kl{$\Nat$-polyregular functions}
    are non-negative.
\end{proof}

@@

\begin{proof}[Proof of \cref{negative-but-npoly:ex}]
    Remark that $P = (X-1)^2 + 1$. Let $f  \colon \Sigma^* \to \Nat$
    that maps $aw \mapsto |w|^2$, and $\varepsilon \to 0$.
    It is clear that $f \in \NPoly$, and that 
    $P$ is \kl{represented} by
    $\ind{|w| \geq 1} \times f + \ind{|w| = 0}$.
\end{proof}

@@

\begin{proof}[Proof of \cref{n-rat-correct:lem}]
    Let $f$ be the \kl{commutative}
    \kl{$\Nat$-rational series}
    with domain $\Sigma \defined \set{a_1, \dots, a_p}$
    that \kl{represents} $P$. Because $f$ has \kl{polynomial growth},
    it is in fact a \kl{$\Nat$-polyregular function}
    (\cref{polynomial-rational-polyreg:fact}).
    Using \cref{n-poly-combinatorics:lem},
    there exists a number $\omega \in \Nat$
    and polynomial $Q$ with \kl{non-negative}
    \kl{maximal monomials} such that
    for all $X_1, \dots, X_p \geq \omega$:
    \begin{equation*}
        f\left(
            \prod_{i = 1}^p a_i^{\omega X_i}
        \right)
        = Q(X_1, \dots, X_p)
        = P(\omega X_1, \dots, \omega X_p) 
        \quad .
    \end{equation*}
    We conclude that $P(\omega X_1, \dots, \omega X_p) = Q(X_1, \dots, X_p)$
    as polynomials,
    and in particular that
    the \kl{maximal monomials} of 
    $P$ are \kl{non-negative}.

    Remark that for every partial valuation $\nu \colon \vec{X} \topartial \Nat$,
    the polynomial $\restr{P}{\nu}$ continues to be represented
    by a \kl{$\Nat$-polyregular function}, namely
    $f$ partially applied to a word. As a consequence,
    the \kl{maximal monomials} of
    $\restr{P}{\nu}$ are also \kl{non-negative}, 
    and
    we have proven that $P \in \CorrectPoly$.
\end{proof}

\begin{fact}
    \label{maxi-monomials-submodular:fact}
    Let $P, Q$ be two polynomials.
    Then, $\MaximalMonomials(P+Q) \hoareleq \MaximalMonomials(P)
    \cup \MaximalMonomials(Q)$.
\end{fact}

\begin{fact}
    \label{translation-maximal:fact}
    For all $P \in \Rel[\vec{X}]$, and $K \in \Nat$,
    \begin{equation*}
        \MaximalMonomials(P) = \MaximalMonomials(\translate{K}(P))
        \quad .
    \end{equation*}
\end{fact}

@@

\begin{proof}[Proof of \cref{decidability-correct:lem}]
    Given $P \in \Rel[\vec{X}]$,
    one applies \cref{derivation-translation:lem}
    to obtain a bound $K$, and then 
    computes $\translate{K}(\restr{P}{\nu})$ for all 
    partial valuations $\nu \colon \vec{X} \topartial \set{0, \dots, K}$.
    Then $P \in \CorrectPoly$ if and only if all those polynomials
    belong to $\Nat[\vec{X}]$, which is decidable.
\end{proof}

@@

\begin{proof}[Proof of \cref{decompose-polynomial:lem}]
    We apply
    \cref{n-poly-combinatorics:lem} to compute the $\omega$.
    Using this $\omega$,
    for any choice of
    $r \defined (r_1, \dots, r_n) \in \set{0, \dots, \omega - 1}$,
    there exists a polynomial $P_r \in \Rel[X_1, \dots, X_p]$
    such that $X_1, \dots, X_p \geq \omega$:
    \begin{equation*}
        f\left(
        a_1^{\omega X_1 + r_1} \cdots a_n ^{\omega X_n + r_n}\right) 
        = P_r(X_1, \dots, X_p) \quad .
    \end{equation*}
    To get the full family of polynomials needed requires
    just to only iterate subsets of letters, and fix the other ones
    to be below the threshold $\omega^2$.
\end{proof}

\begin{proof}[Proof of \cref{decidable-n-poly:thm}]
    \textbf{TODO: we need to talk about the number of variables hehe}
    Let us prove that
    \cref{f-npoly-combi:item}
    implies \cref{f-combinatorial:item}. 
    We compute the $\omega$ by applying 
    \cref{n-poly-combinatorics:lem}, and then
    notice that the polynomial $P$ obtained is
    by definition a \kl{$\Nat$-rational polynomial}.

    For the converse implication, we use \cref{decompose-polynomial:lem} to
    effectively compute a polynomial representation of $f$. Using the
    assumption, all of these polynomials are actually \kl{$\Nat$-rational
    polynomials}. Let us now construct a new way to compute $f$, by first
    computing the \kl{$\omega$-type} of the input (which is possible in
    $\MSO$), and then computing the suitable \kl{$\Nat$-rational polynomial},
    which is possible in $\NPoly$.
\end{proof}

\begin{proof}[Proof of \cref{zsf-npoly-nsf:thm}]
    We proceed by induction on the size of the alphabet $\card{\Sigma}$.

    \textbf{Base Case.} When $\Sigma$ is empty, $\Sigma^*$ contains
    only the empty word $\varepsilon$, and because $f$ is a
    \kl{$\Nat$-polyregular function}, $f(\varepsilon) \in \Nat$.
    Note that $f$ is therefore a constant function, equal to a natural
    number, hence belongs to $\NSF$.

    \textbf{Induction.}
    Because $f$ is \kl{ultimately polynomial},
    there exists $N_0 \in \Nat$, 
    and $P \in \Rel[\seqof{X_a}{a \in \Sigma}]$ 
    such that
    \begin{equation*}
        f\left( \prod_{a \in \Sigma} a^{X_a} \right)
        = 
        P(\seqof{X_a}{a \in \Sigma})
        \quad
        \text{ when }
        \forall a \in \Sigma, X_a \geq N_0
        \quad .
    \end{equation*}
    Furthermore, leveraging \cref{decompose-polynomial:lem},
    we obtain $\omega \in \Nat$
    and a family of polynomials 
    $\seqof{P_{t}}{t \in \ModuloTypes[\omega]^\Sigma}$
    that represent $f$.
    For all $t \in \ModuloTypes[\omega]^\Sigma$
    that have threshold greater than $\omega^2$,
    we conclude that $P_t$ and $P$ coincide (up to translation) over an infinite grid,
    hence we conclude that
    if all letters appear at least $N_1 \defined \max (\omega^2, N_0)$ times,
    then $f$ computes $P$. But this also shows that
    $\translate{N_1}(P)$ is a \kl{$\Nat$-rational polynomial}, hence is
    \kl{represented} by a function $f_1 \in \NSF[k]$.

    To compute $f$ on words where some letter appears fewer than $N_1$ times,
    we partially apply $f$ to obtain a function $g$ over a restricted input
    alphabet. Using the induction hypothesis, $g$ belongs to $\NSF[k]$, and we
    can therefore (effectively) compute a representation of $f$ in $\NSF[k]$.

    For the effective part of the theorem, 
    we know from \cite[Theorem V.13]{LOPEZ23b}, it is decidable whether
    $f \in \ZSF[k]$ and from \cref{decidable-n-poly:thm}
    it is decidable whether $f \in \NPoly$.
    Furthermore, one can enumerate functions in $\NSF[k]$,
    and check equivalence to effectively represent $f$ as a
    \kl{star-free $\Nat$-polyregular function}.
\end{proof}

@@

\begin{fact}
    \label{unique-res-transducer:fact}
    Let $f \colon \Sigma^* \to \Nat$ and $k \in \Nat$.
    Then $f$ has at most one \kl{$k$-residual transducer}.
\end{fact}
\begin{proof}[Proof of \cref{unique-res-transducer:fact}]
    Let $\aTransd_1$ and $\aTransd_2$ be two
    \kl{$k$-residual transducers} for $f$.
    The two initial states must be $\varepsilon$.
    Let us prove by induction on $u \in \Sigma^*$ that
    $\delta_1(\varepsilon, u) = \delta_2(\varepsilon, u)$
    and that $Q_1$ equals $Q_2$ over prefixes of $u$.
    This will prove that 
    $Q_1 = Q_2$, hence that $\aTransd_1 = \aTransd_2$.

    Let $u \in \Sigma^* \cap Q_1 \cap Q_2$ and $a \in \Sigma$, $v_1 \in Q_1$ be
    defined as $v \defined \delta_1(u,a)$, and $v_2 \defined \delta_2(u,a)$.
    Remark that by induction hypothesis, for all $v \prefle u$, $v \in Q_1 \cap
    Q_2$. If $\delta_1(u,a) = Q_1$, it means that for all $v \in Q_2$ such that
    $v \prefle ua$, we have $\neg( v \resleq{f}{k} ua )$. The only possible
    transition in $\aTransd_2$ is therefore $\delta_2(u,a) = ua$, and $ua \in
    Q_2$. Similarly, if $\delta_1(u,a) \prefleq u$, then $\delta_2(u,a) =
    \delta_1(u,a)$ by definition of $\delta_2$ as a maximum.
\end{proof}

\begin{fact}
    \label{q-o-prefix-cool:fact}
    Let $f \colon \Sigma^* \to \Nat$ and $k \in \Nat$.
    At each step of the \texttt{while loop}
    of \cref{residual:algo}, the sets
    $Q$ and $O$ are such that
    \begin{enumerate}
        \item $Q \cup O$ is a \kl{downwards closed} subset of 
            $\Sigma^*$ for $\prefleq$;
        \item elements in $O$ are pairwise incomparable
            for $\prefleq$, and are maximal
            for $\prefleq$ inside $Q \cup O$.
    \end{enumerate}
\end{fact}
\begin{proof}[Proof of
    \cref{q-o-prefix-cool:fact}]
    Let us write $Q_i$ and $O_i$ for the value of the variables
    $Q$ and $O$ at step $i$ of the \texttt{while loop}.
    We prove the desired property by induction on $i$.

@@

    If the \texttt{else} branch was taken, then there exists $u \in O_i$ such
    that $Q_{i+1} = Q_i \cup \set{u}$ and $O_{i+1} = O_i \setminus \set{ u }
    \cup \setof{ ua }{ a \in \Sigma}$. We conclude that $Q_{i+1} \cup O_{i+1} =
    Q_i \cup O_i \cup \setof{ ua }{a \in \Sigma}$ continues to be \kl{downwards
    closed} for $\prefleq$. Let $v \in Q_{i+1} \cup O_{i+1}$ be such that $ua
    \prefleq v$ for some $a \in \Sigma$. Then $u \prefleq v$, and $u = v$ since
    $u$ was a maximal element. As a consequence, $ua$ is a maximal element for
    all $a \in \Sigma$. Assume by contradiction that $ua$ is comparable with
    some $v \in O_{i+1}$ with $ua \neq v$, it cannot be that $ua \prefleq v$ by
    the above argument, and if $v \prefleq ua$ with $v \neq ua$, then $v
    \prefleq u$ and $u = v$, which is absurd since $v \not \in O{i+1}$.
    We have concluded that $O_{i+1}$ continues to have pairwise incomparable
    elements.
\end{proof}

\begin{proof}[Proof of \cref{correct-residual:lemma}]
    Because of \cref{q-o-prefix-cool:fact},
    we already know that $q_0 = \varepsilon$,
    $Q$ is a \kl{downwards closed} subset of $\Sigma^*$
    for $\prefleq$, 
    that every state of $Q$ is accessible from $q_0$.
    Notice that at every step,
    $\lambda(u,a)$ is defined as
    $\app{f}{ua} - \app{f}{\delta(u,a)}$.
    Finally, since $Q \cup O$ is a \kl{downwards closed} subset of $\Sigma^*$
    at every step,
    we have that at step $i$,
    for all $ua \in O_i$,
    $\setof{w \in Q}{w \prefleq ua} = \setof{w \in Q_i}{ w \prefleq ua}$,
    which proves that the maximum considered in the algorithm
    is indeed computing correctly.
\end{proof}

\begin{proof}[Proof of \cref{wqo-implies-termination:lemma}]
    Assume towards a contradiction that
    \cref{residual:algo} does not terminate.
    Then, the \texttt{else} branch in the \texttt{while loop}
    must be taken infinitely often.
    This means that the set $Q$ of states grows arbitrarily large.

    Let us write $\seqof{Q_i}{i \in \Nat}$ for the set of states $Q$ at step
    $i$ of the execution of \cref{residual:algo}. Applying
    \cref{q-o-prefix-cool:fact}, we know that for all $i \in \Nat$, $Q_i$ is
    \kl{downwards closed} for $\prefleq$. Let us write $Q_\infty \defined
    \bigcup_{i \in \Nat} Q_i$. The set $Q_\infty$ is infinite, and is
    \kl{downwards closed} for $\prefleq$. As a consequence, it is an infinite
    tree with a finite branching (at most $\card{\Sigma}$), and has an infinite
    branch $\seqof{u_j}{j \in \Nat}$ thanks to König' s lemma.

@@

    As a consequence, if $\MSO^q(u) \leq \MSO^q(v)$, then 
    for all $w \in \Sigma^*$: 
    \begin{align*}
        & f(vw) - f(uw) \\
        &= 
        \sum_{i = 1}^n m_i
        \left[
            \vcount{\phi_i(\vec{x})} (vw) -
            \vcount{\phi_i(\vec{x})} (uw)
        \right] \\
        &= 
        \sum_{i = 1}^n
        m_i
            \sum_{\phi_i \in t(\vec{x})}
        \left[
            \vcount{t(\vec{x})}(vw)
            -
            \vcount{t(\vec{x})}(uw)
        \right] \\
        &= 
        \sum_{i = 1}^n
        m_i
        \sum_{1 \leq j \leq N_0}
        \sum_{\phi_i \in t_l^j(\vec{y}) \odot t_r^j(\vec{z})}
        \underbrace{
        \left[
            \vcount{t_r^j(\vec{y})}(v)
            -
            \vcount{t_r^j(\vec{y})}(u)
        \right] 
    }_{ \in \Nat }
            \times 
            \vcount{t_l^j(\vec{z})}(w)
    \end{align*}

@@

\begin{proof}[Proof of \cref{aperiodic-iff-residual:lem}]
    It is clear that $\NSF \subseteq \ZSF$, and known that if $f \in \ZSF$
    then it is \kl{ultimately polynomial}. Furthermore, if the \kl{$0$-residual
    transducer} of $f$ is \kl{counter-free}, then $f \in \NSF$
    using \cref{transducer-nsf-npoly:lemma}.

    Assume that $f$ is \kl{ultimately polynomial}, let us prove that the
    \kl{$0$-residual transducer} of $f$ is \kl{counter-free}. 
    Note that because $f \in \NPoly[0]$,
    $u \resleq{f}{0} v$ if and only if $\app{f}{u} = \app{f}{v}$.
    In particular, in a \kl{$0$-residual transducer} of $f$,  two states that
    represent the same residual must be incomparable for the prefix relation.

    Let $(q,w^n)$ be
    a counter with $n \geq 1$. This means that $\delta(q, w^n) = q$ in the
    automaton, and implies that $q \resleq{f}{0} qw^n$, hence that $\app{f}{q}
    = \app{f}{qw^n}$. Because $f$ is \kl{ultimately polynomial},
    we know that $\app{f}{qw^n} = \app{f}{qw^{n+1}}$,
    hence that $\app{f}{qw} = \app{f}{q}$.

    Let us write $t \defined \delta(q,w) = \delta(q,w^{n+1})$. We know that
    $\app{f}{q} = \app{f}{t}$. Assume by contradiction that $t$ and $q$ are
    incomparable for the prefix relation. Let us split $w = w_1 w_2$ where
    $w_1$ is the shortest prefix of $w$ such that $s_0 \defined \delta(q,w_1)$
    is an ancestor of $q$ and of $t$ for the prefix relation, it must exist
    because $\delta(q,w_1 w_2) = t$.

    Now, consider $s_1 \defined \delta(t, w_1)$. Assume by contradiction that
    $s_0$ is not comparable with $s_1$ for the prefix relation. Then, consider
    the smallest prefix $v$ of $w_1$ such that $\delta(t, v)$ is a strict
    prefix of $s_0$. It must exist, otherwise $s_0$ is always a prefix of
    $s_1$. Because $\app{f}{t} = \app{f}{q}$, we conclude that $\app{f}{tv} =
    \app{f}{qv}$. However, this contradicts the minimality of $w_1$, since
    $\delta(t,v)$ is an ancestor of $q$ and $t$.

    We have proven that $s_0$ and $s_1$ are comparable, hence they are equal,
    since $\app{f}{s_1} = \app{f}{t w_1} = \app{f}{q w_1} = \app{f}{s_0}$.
    Finally, we have proven that $\delta(q, w_1) = s_0$, $\delta(s_0, w_2) =
    t$, and $\delta(s_0, w_2) = \delta(t, w_1w_2) = q$ which is absurd.
    As a consequence $t$ and $q$ were comparable for the prefix relation,
    hence equal, and therefore $\delta(q, w) = q$.
\end{proof}

\begin{proof}[Proof sketch of \cref{sf-no-periods-on-sequences:lemma}]
    Let $t$ be a first order type,
    the map $g_t \colon X \mapsto \vcount{t}(uw^X)$ is a
    \kl{commutative} \kl{star-free $\Nat$-polyregular function},
    hence ultimately a polynomial, and therefore,
    translated enough, it is a $\Nat[X]$ polynomial.
    But this means that $g_t(X+1) - g_t(X) \in \Nat[X]$ for a large
    enough $X$, which is precisely what suffices.
    Now, we can do this for every first order type $t$
    with fewer variables and lower quantifier rank than those
    needed to define $f$, and conclude.
\end{proof}

### 2024-02-12 parts/conclusion.tex
big big rewrite

%! TeX program = xelatex
%! lang = en-US
\section{Conclusion}
\label{sec:ccl}

@@

\begin{center}
    \begin{tabular}{c|cc}
        \toprule
        & \textbf{Commutative} & \textbf{General} \\
        \midrule
        \textbf{Commutative}
        and \textbf{Invertible} & commutative $\ZPoly$ & $\ZPoly$ \\
        \textbf{Commutative} & commutative $\NPoly$ & $\NPoly$ \\
        \textbf{General} & ? & polyregular 
    \end{tabular}
\end{center}

>>>>

%! TeX program = xelatex
%! lang = en-US
\section{Outlook}
\label{sec:ccl}

While \kl{$k$-residual transducers} will provide canonical objects associated
to functions in $\NPoly$, their effective construction requires the ability to
decide $\NPoly$ inside $\ZPoly$, which is not currently known (except in the
commutative case). Furthermore, the correspondence between \kl{counters} in
this automaton and \kl{star-free $\Nat$-polyregular functions} is non-trivial,
and cannot be used \emph{as is} to decide membership in $\NSF$
based on this analysis of the \kl{$k$-residual transducer}.

Let us prove that the \kl{residual transducer} is enough to decide aperiodicity
in the case of functions in $\NPoly[0]$, that is, with finite image. In
particular, this generalizes the result of \cite{LOPEZ23b}, by proving that
$\NSF[0] = \NPoly[0] \cap \ZSF$. Even though the proposed model is canonical,
and allows to effectively decide aperiodicity for \kl{$\Nat$-polyregular
functions} of growth rate $0$ (that is, with bounded image), it is not clear
yet if it can be leveraged to decide aperiodicity for \kl{$\Nat$-polyregular
functions} of higher growth rates.

\begin{lemma}
    \label{aperiodic-iff-residual:lem}
    Let $f \in \NPoly[0]$, the following are equivalent:
    \begin{enumerate}
        \item $f \in \NSF$,
        \item $f \in \ZSF$.
        \item $f$ is \kl{ultimately polynomial},
        \item The \kl{$0$-residual transducer} of $f$ is \kl{counter-free},
    \end{enumerate}
\end{lemma}

\textbf{TODO:} one may think that we can ask that $u \resleq{f}{k} u w^n$
implies $u \resleq{f}{k} uw$, but it is false, as the following example will
demonstrate. This is why the definition of counter is more involved in our
case.

\begin{lemma}
    \label{sf-no-periods-on-sequences:lemma}
    Let $f \colon \Sigma^* \to \Nat$ be a \kl{star-free $\Nat$-polyregular
    function}. Then,
    for all $u, w \in \Sigma^*$,
    there exists $n \in \Nat$ such that
    $u w^n \resleq{f}{k} u w^{n+1}$.
\end{lemma}

\begin{theorem}
    The following are equivalent for \kl{commutative}.
    \begin{itemize}
        \item $f$ is in $\NSF$,
        \item The residual automata of $f$ has no counters
    \end{itemize}
\end{theorem}

@@

 This suggests the
following conjecture in the non-commutative case:

\begin{conjecture}
    Let $k \in \Nat$,
    and $f \in \ZPoly[k]$. Then, $f \in \NPoly[k]$
    if and only if $f$ is \kl{combinatorial}, which is decidable,
    and an effective conversion procedure exists.
\end{conjecture}
